Stan Introduction

### 2019-08-26 Code/Stan-introduction.Rmd
adding code chunks for stan code
added chains code and explanation, and other minor edits.

***
data {

@@

***

@@

***
data {  
  int N;  
  real Y[N];  
}

@@

model {  
  for(i in 1:N)  
   Y[i] ~ normal(mu, sigma);  
  mu ~ normal(1.7, 0.3)  
  sigma ~ cauchy(0, 1);  
}

***

@@

```{r}
mymodeldata = list(
  N = 100,
  Y = rnorm(N, 1.6, 0.2)
)
fit = stan(
  file = here::here("Code", "my_stan_model.stan"),
  data = mymodeldata,
  seed = 42
)
```

@@

Now the object fit contains the results of running our data through our model. Let's check them out.
```{r}
print(fit)
```
Here we can see the estimated means for our parameters and diagnose our model for issues with convergence. You will learn more abot that later, but the model seems to have run correctly here. We can also look at the posterior with the following code.
```{r}
parametermu = extract(fit)$mu
hist(parametermu)
```
Here we have a posterior distribtution for mu. Since our prior and simulated data were very similar, we see a very tight distribution, but again, try it with different means for both the sample and the prior, and see how the data interracts.

>>>>

```
data {

@@

@@

```
data {  
  int N;  
  real Y[N];  
}

@@

model {  
//This for loop is our likelyhood
  for(i in 1:N)  
   Y[i] ~ normal(mu, sigma);  
//These are our priors
  mu ~ normal(1.7, 0.3)  
  sigma ~ cauchy(0, 1);  
}

@@

```{r}
mymodeldata = list(
  N = 100,
  Y = rnorm(N, 1.6, 0.2)
)
fit = stan(
  file = here::here("Code", "my_stan_model.stan"),
  data = mymodeldata,
  seed = 42,
  chains = 4
)
```

@@

The chains code is helpful for making your code run more efficiently and checking for accuracy. We will go into more detail later, but for now you should run 4 chains, assuming your computer has 4 cores. If your not sure how many cores you have, you can usually find that information in your computer settings, or you can run this code.

```
install.packages("parellel")
library(parallel)
detectCores()
```
R will detect the number of cores on your PC.

Now the object 'fit' contains the results of running our data through our model. Let's check them out.
```{r}
print(fit)
```
Here we can see the estimated means for our parameters and diagnose our model for issues with convergence. You will learn more abot that later, but the model seems to have run correctly here. We can also look at the posterior with the following code.
```{r}
parametermu = extract(fit)$mu
hist(parametermu)
```
Here we have a posterior distribtution for mu. Since our prior and simulated data were very similar, we see a very tight distribution, but again, try it with different means for both the sample and the prior, and see how the data interracts.
